---
title: "Fitbit Analysis"
author: "Sara Kim"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

```{r}
fitbit = readxl::read_xlsx("./data/final.xlsx") %>%
  janitor::clean_names()
```

Goal:
The primary exposure and outcome will be the mean number of steps/day and the risk of spontaneous preterm delivery (<37 weeks), respectively. We will assess if the mean number of steps/day will differ between women that spontaneously deliver preterm versus those that deliver at term. 

## EDA

```{r}
fitbit[,c(3:5)] %>%
 summary()
```

Comparison:
United States: 5,117 steps. This is about 2.5 miles or about 4 kilometers each day.
Switzerland: 9,650 steps. This is about 4.8 miles or 8 kilometers each day.
Japan: 7,168 steps. This is about 3.5 miles or 6 kilometers each day.

We can notice large variance in steps and higher steps than us mean.

```{r}
ggplot(fitbit, aes(x=step_total)) +
  geom_histogram(color ="black", fill="white")
```

```{r}
fitbit %>%
  group_by(term) %>%
  summarize(median_steps = median(step_total),
         mean_steps = mean(step_total),
         sd_steps = sd(step_total)) %>%
  knitr::kable()
```
Non-preterm has a larger variance but both groups have very similar number of steps.

## Longitudinal Effect

```{r}
fitbit = fitbit %>%
  mutate(
      day_ctg = case_when(
      day < 92 ~ "-91",  
      day %in% 92:140 ~ "92-140",
      day %in% 141:189 ~ "141-189",
      day %in% 190:238 ~ "190-238",
      day %in% 239:288 ~ "239-288"
    ),
    day_ctg = fct_relevel(day_ctg, "-91")) 
```

```{r}
fitbit %>%
  ggplot(aes(x = day_ctg, y = step_total, group = id)) + geom_line(aes(alpha = 0.2)) + stat_smooth(aes(group = 1), method = "lm", se = FALSE) +
    stat_summary(aes(group = 1), geom = "point", color = "green", fun.y = mean, shape = 17, size = 3) +
    facet_grid(. ~ term)
```

I windowed the gestational age (day) into 5 windows. It does not seem like there is a difference in the steps based on the gestational age.

## Median Steps per Day

```{r}
fitbit_median = fitbit %>%
  group_by(id, term) %>%
  summarise(median_step = median(step_total),
            mean_step = mean(step_total),
            ga_delivery = mean(ga_delivery))

fitbit_median %>%
  group_by(term) %>%
  summarise(n = n(),
            percent = n()/nrow(fitbit_median)) %>%
  knitr::kable()

ggplot(fitbit_median, aes(term,median_step, fill=term)) + 
    geom_boxplot()

fitbit_median %>%
  group_by(term) %>%
  summarise(n = n(),
            median_step= median(median_step))
```

Looking at the median steps for each indivisual, the median of the median steps seems similar between the two groups. The median value of the pre-term birth cases is actually slightly higher (`r 7765 - 6971` steps).

```{r}
term_fct = as.factor(ifelse(fitbit_median$term == "TERM", 0, 1))
fitbit_median["term_fct"] = term_fct

fitbit_median %>%
  ggplot(aes(x=median_step, y=term_fct)) +
  geom_point()
```

## logistic regression

```{r}
glm <- glm(term_fct ~ median_step, family=binomial(link='logit'), data = fitbit_median) 

glm  %>% 
  broom::tidy() %>%
  knitr::kable(digits = 3)
```

there is no statistically significant evidence that there is a linear relationship between log odds of pre-term birth and median steps per day.

## Relationship Between Delivery Date and Median Steps Per Day

```{r}
fitbit_median %>%
  ggplot(aes(x=median_step, y= ga_delivery, color = term)) +
  geom_point()
```

```{r}
res <- cor.test(fitbit_median$median_step, fitbit_median$ga_delivery, 
                    method = "pearson")
res
res$p.value
```

